### 2025-10-21 reports/draft_results.qmd
heatmap revise

---
title: "DOM bridges and barriers: systematic review and survey"
author: "Emily Nazario | UC Santa Cruz | enazario'@'ucsc.edu"
date: "`r Sys.Date()`"
format:
 html: 
  self-contained: true
editor: visual
toc: TRUE
toc-title: "On this page"
theme: yeti
fontcolor: "#134f5c"
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(tidyquant)
library(here)
library(kableExtra)
library(waffle)
library(extrafont)
library(hrbrthemes)
library(patchwork)
library(ggalluvial)
library(ggrepel)
library(forcats)

#lit data
code_freq <- read.csv(here("data/code_freq_lit.csv"))
code_vars <- read.csv(here("data/code_w_vars_lit.csv"))
keywords <- read.csv(here("data/keywords.csv"))
code_rel <- read.csv(here("data/code_rel.csv"))
code_rel2 <- read.csv(here("data/br_ba_code_lit.csv"))

#survey data
code_vars_s <- read.csv(here("data/surveys/code_w_vars.csv"))
code_freq_s <- read.csv(here("data/surveys/code_freq.csv"))
code_rel_s <- read.csv(here("data/surveys/br_ba_code_survey.csv"))

#change codes to new names 
  #code_freq
    #sp br  
code_freq$code_cat <- replace(code_freq$code_cat, code_freq$code_cat == "Species bridge", "Ecological bridge")
code_freq$code_cat <- replace(code_freq$code_cat, code_freq$code_cat == "Species barrier", "Ecological barrier")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Limited Prey Resources", "Target species is a prey specialist")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Fine Scale management required", "Fine-scale management appropriate for target species")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Migratory and/or dynamic range", "Highly mobile target species")
code_freq$code_name[code_freq$code_cat == "Ecological bridge" & code_freq$code_name == "Climate change"] <- "Expected climate change induced range shifts"
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Known Species Nat History", "Life history information available for target species")

    #data br
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Advances in Tech", "Access to advanced technology and techniques")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Scale appropriate for system", "Management scale aligns with system needs")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Availability for species", "Target species data availability")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Stakeholder/Anthropogenic Data Access", "Resource use data availability")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Enviro Access", "Abiotic data availability")

    #soc br
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Open Communication or Co-Mngmt in place", "Co-management and communication")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Pre-existing regulation & management institutions", "Pre-existing regulation and management institutions")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "DOM implementation feasible (Socio-econ)", "Feasibility")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Resource Availability (economical)", "Resource availability")

  #sp ba 
code_freq$code_cat[code_freq$code_name == "Model Abuse"] <- "Data barrier"
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Switch to multi-species mngmnt", "Need for multispecies management")

  #data ba
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Gaps and Accuracy", "Gaps and inaccuracy")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Model Abuse", "Model abuse")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Mismatch in scale", "Management scales do not align with system needs")
code_freq$code_cat[code_freq$code_name == "Climate change"] <- "Data barrier"
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Climate change", "Climate change adding uncertainty to target species habitat use")

  #soc ba
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Unintended consequences", "Indirect socioeconomic consequences")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Communication & Distribution", "Poor communication and outreach")
code_freq$code_name <- replace(code_freq$code_name, code_freq$code_name == "Recipient participation", "Poor program implementation")

theme_ms <- function(){
  theme_minimal() %+replace%    
    theme(axis.text = element_text(size = 14, color = "black"), 
          axis.title = element_text(size = 16, color = "black"), 
          strip.text = element_text(size = 13, color = "black"))
}

```

## Questions

**Which themes appear in both sources?**

**What new information have surveys added beyond what was already written in the literature?**

**How do themes vary between fishing vs. shipping contexts?**

**How do themes vary with positionality of the respondent (surveys only)?**

## Result 1: Bridges and barriers according to review and survey

```{r}
#| echo: false
#| message: false
#| warning: false
#| include: false
#| fig-width: 15
#| fig-height: 15

#calc props for sankey_df csv (fine scale codes are raw totals from code_freq file)
sum_codes <- sum(code_freq$total) #710 codes
code_freq %>% 
  group_by(bridge_barr) %>% 
  summarise(totals = sum(total))

code_freq %>% 
  group_by(code_cat) %>% 
  summarise(totals = sum(total))

#load sankey df
sankey_df <- read.csv(here("data/sankey_freq.csv"))

sankey_df$code_cat <- factor(sankey_df$code_cat, levels = c("Ecological barrier", "Data barrier", "Social barrier", "Data bridge", "Ecological bridge", "Social bridge"))

sankey_df <- sankey_df %>%
mutate(sum_codes = sum(total), 
      perc_codes = (total/sum_codes)*100, 
      round_perc = round(perc_codes, 0))

ggplot(data = sankey_df,
       aes(axis1 = fct_reorder(bridge_barr, -perc_codes), axis2 = fct_reorder(code_cat, -perc_codes), axis3 = fct_reorder(code_name, -perc_codes), y = perc_codes)) +
  geom_alluvium(aes(fill = code_cat, color = code_cat), size = 1.5, alpha = 0.8, width = 0.05) +
  geom_stratum(size = 1.5, color = "grey60", 
              width = c(1/12,1/12,1/3,1/3,1/3,1/3,1/3,1/3,1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18, 1/18)) +
  geom_text(stat = "stratum", 
    aes(label = after_stat(stratum)),
    angle = c(90,90,0,0,0,0,0,0,0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), 
    size = c(7,7,5,5,5,5,5,5,4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 4), 
    nudge_y = c(0,0,0,0,0,0,0,0,0, #poor communication and outreach
                0.5, #model abuse
                1, #Discontinuity
                1.5, #indirect socioeconomic consequences
                1.8, #target species is a prey specialist
                2, #resource availability
                2, #climate change adding uncertainty
                2, #data transparency
                2, #need for multi-species management
                2, #distinct habitat preference
                2, #fine scale management
                1.5, #management scale system needs
                2, #management scale aligns
                0.8, #feasibility
                0,0,0,0,0,0,0,0,0,0,0,0,0,0), 
    nudge_x = c(-0.01,-0.01,0,0,0,0,0,0,0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05,   0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05), 
    hjust = c(0.5, 0.5, 0.5,0.5,0.5,0.5,0.5,0.5,0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)) +

  xlim(0.89, 3.6) +
  ylim(-1, 100)+
  labs(y = "Proportion of all codes (%)", x = "") +
  theme_minimal()+
  theme(axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16), 
        legend.position = "none", 
        axis.text.x = element_blank(), 
        panel.grid.major = element_blank())+
  scale_fill_manual(values = c("#F48136", "#F46036", "#F44336", "#CFE2F3", "#C5E0B4", "#45818E"))+
  scale_color_manual(values = c("#F48136", "#F46036", "#F44336", "#CFE2F3", "#C5E0B4", "#45818E"))
```

### Sunburst figure

Visualizes each of the themes (Data, Ecological, Social, and Implementation) identified in the literature and surveys. Also visualizes the sub-themes within each theme bin. Literature themes are on top while surveys are on the bottom. Cooler colors indicate bridges while warmer ones represent barriers. The small text/poor resolution should be resolved when the figure is viewed as a pdf!

![](images/sunburst.png){fig-align="center" width="100%"}

### Table

May include in the supplemental, but could be redundant to the sunburst diagram.

::: panel-tabset
#### Review: Bridge themes

```{r}
#| echo: false
#| message: false
#| warning: false

bri_codes <- code_freq %>% 
  filter(bridge_barr == "Bridge") %>% 
  subset(select = -c(total, bridge_barr))

colnames(bri_codes) <- c("Code category", "Code label")

bri_codes %>%
  kbl() %>%
  kable_styling()
```

#### Review: Barrier themes

```{r}
#| echo: false
#| message: false
#| warning: false

bar_codes <- code_freq %>% 
  filter(bridge_barr == "Barrier") %>% 
  subset(select = -c(total, bridge_barr))

colnames(bar_codes) <- c("Code category", "Code label")

bar_codes %>%
  kbl() %>%
  kable_styling()
```

#### Survey: Bridge themes

```{r}
#| echo: false
#| message: false
#| warning: false

bri_codes_s <- code_freq_s %>% 
  filter(bridge_barr == "Bridge") %>% 
  subset(select = -c(total, bridge_barr))

colnames(bri_codes_s) <- c("Code category", "Code label")

bri_codes_s %>%
  kbl() %>%
  kable_styling()

```

#### Survey: Barrier themes

```{r}
#| echo: false
#| message: false
#| warning: false

bar_codes_s <- code_freq_s %>% 
  filter(bridge_barr == "Barrier") %>% 
  subset(select = -c(total, bridge_barr))

colnames(bar_codes_s) <- c("Code category", "Code label")

bar_codes_s %>%
  kbl() %>%
  kable_styling()

```
:::

## Result 2: Codes by review and survey

### Result 2.1: Frequencies according to review and survey

Code appearance counts across all bridge **OR** barrier codes. The values represent counts of themes or sub-themes.

::: panel-tabset
### Literature

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center
#| include: FALSE

  code_sum <- code_freq %>%
  #mutate(n_total = sum(total)) %>%
  #group_by(code_cat) %>% 
  #mutate(n_subtotal = sum(total)) %>%
  #summarise(value = n_subtotal/n_total) %>%
  rename("br_ba" = code_cat) %>%
  # mutate(percent_label = scales::percent(value, accuracy = 1), 
  #        percent = if_else(br_ba == "Data barrier" | br_ba == "Ecological barrier" | br_ba == "Social barrier", value, -value),
  #        code_level = if_else(br_ba %in% c("Ecological barrier", "Ecological bridge"), "Ecological bridge or barrier", 
  #                             if_else(br_ba %in% c("Data barrier", "Data bridge"), "Data bridge or barrier",
  #                             if_else(br_ba %in% c("Social barrier", "Social bridge"), "Social bridge or barrier", NA))), 
  #        code_brba = if_else(br_ba %in% c("Data bridge", "Ecological bridge", "Social bridge"), "bridge", "barrier")) %>%
  mutate(percent = if_else(br_ba == "Data barrier" | br_ba == "Ecological barrier" | br_ba == "Social barrier", -total, total),
         code_level = if_else(br_ba %in% c("Ecological barrier", "Ecological bridge"), "Ecological", 
                              if_else(br_ba %in% c("Data barrier", "Data bridge"), "Data",
                              if_else(br_ba %in% c("Social barrier", "Social bridge"), "Social", NA))), 
         code_brba = if_else(br_ba %in% c("Data bridge", "Ecological bridge", "Social bridge"), "bridge", "barrier")) %>%
  group_by(code_brba, code_level) %>%
  summarize(value = sum(percent)) %>%
  mutate(value_label = ifelse(code_brba == "barrier", value*-1, value)) %>%
  distinct()

code_sum$code_level <- factor(code_sum$code_level, levels = c("Social", "Ecological", "Data"))
code_sum$code_brba <- factor(code_sum$code_brba, levels = c("barrier", "bridge"))

ggplot(code_sum, aes(x = code_level, 
             y = value,
             fill = code_brba)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = value_label),
            position = position_stack(vjust = 0.5),
            color = "black", 
            size = 5) +
  #ylim(-0.4,0.7)+
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"), labels = c("Barrier", "Bridge")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x =element_blank(),
        axis.text.y = element_text(size = 20, color = "black"),
        panel.grid = element_blank(),
        legend.position = c(-0.05, 1),
        legend.text = element_text(size = 16, color = "black"),
        strip.text.y.right = element_text(angle = 0, size = 16)) +
  geom_hline(yintercept = 0.0, linewidth = 1) + 
  labs(fill = "")+ 
  guides(fill = guide_legend(nrow = 1))

#ggsave("figures/ms_figs/all_freq/all_freq.png", width = 8, height = 6, units = ("in"))
#ggsave("figures/slides_svg/freq_lit.svg", width = 8, height = 6, units = ("in"))
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center

#reformat data
code_sum_br <- code_freq %>%
  filter(bridge_barr == "Bridge") %>%
  #mutate(n_total = sum(total)) %>%
  #group_by(code_cat, code_name) %>%
  #summarise(value = total/n_total) %>% #percentages represent total number of bridge or barriers codes that code represents
  mutate(#percent = scales::percent(value, accuracy = 1), 
         code_name = factor(code_name)) %>%
  ungroup() %>%
  group_by(code_cat) %>%
  arrange(code_name, desc(total))

#plot
br_freq <- ggplot(code_sum_br, aes(x = total, y = reorder(code_name, total))) + 
  geom_point(size = 5, color = "lightskyblue3") + 
  # scale_x_continuous(limits = c(0, 30), 
  #                    breaks = seq(0, 30, 5), 
  #                    name = "Percent appearance (%)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  scale_x_continuous(limits = c(0, 75), 
                  #breaks = seq(0, 30, 5), 
                  name = "Sub-theme count (#)")+
  scale_y_discrete(position = "right") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 0), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

#ggsave("figures/ms_figs/all_freq/br_freq.png", width = 7, height = 8, units = c("in"))

```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center

#reformat data
code_sum_ba <- code_freq %>%
  filter(bridge_barr == "Barrier") %>%
  #mutate(n_total = sum(total)) %>%
  #group_by(code_cat, code_name) %>%
  #summarise(value = total/n_total) %>% #percentages represent total number of bridge or barriers codes that code represents
  mutate(#percent = scales::percent(value, accuracy = 1), 
         code_name = factor(code_name)) %>%
  ungroup() %>%
  group_by(code_cat) %>%
  arrange(code_name, desc(total))

#plot
ba_freq <- ggplot(code_sum_ba, aes(x = total, y = reorder(code_name, total))) + 
  geom_point(size = 5, color = "goldenrod") + 
  scale_x_reverse(limits = c(75, 0), 
                     breaks = seq(0,60, 20), 
                     name = "Sub-theme count (#)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 1), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

#ggsave("figures/ms_figs/all_freq/ba_freq.png", width = 7.5, height = 8, units = c("in"))

```

![](images/br_ba_freq.png){fig-align="center" width="10in", height = "12in"}

### Surveys

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center

#theme bridge/barrier
 code_sum_s <- code_freq_s %>%
  rename("br_ba" = code_cat) %>%
  mutate(percent = if_else(br_ba == "Data barrier" | br_ba == "Ecological barrier" | br_ba == "Social barrier" | br_ba == "Implementation barrier", -total, total),
         code_level = if_else(br_ba %in% c("Ecological barrier", "Ecological bridge"), "Ecological", 
                              if_else(br_ba %in% c("Data barrier", "Data bridge"), "Data",
                              if_else(br_ba %in% c("Implementation barrier", "Implementation bridge"), "Implementation",
                              if_else(br_ba %in% c("Social barrier", "Social bridge"), "Social", NA)))),
         code_brba = if_else(br_ba %in% c("Data bridge", "Ecological bridge", "Social bridge", "Implementation bridge"), "bridge", "barrier")) %>%
  group_by(code_brba, code_level) %>%
  summarize(value = sum(percent)) %>%
  mutate(value_label = ifelse(code_brba == "barrier", value*-1, value)) %>%
  distinct()

code_sum_s$code_level <- factor(code_sum_s$code_level, levels = c("Implementation", "Social", "Ecological", "Data"))
code_sum_s$code_brba <- factor(code_sum_s$code_brba, levels = c("barrier", "bridge"))

all_freq_survey <- ggplot(code_sum_s, aes(x = code_level, 
             y = value,
             fill = code_brba)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = value_label),
            position = position_stack(vjust = 0.5),
            color = "black", 
            size = 5) +
  #ylim(-0.55,0.7)+
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"), labels = c("Barrier", "Bridge")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x =element_blank(),
        axis.text.y = element_text(size = 20, color = "black"),
        panel.grid = element_blank(),
        legend.position = c(-0.05, 1),
        legend.text = element_text(size = 16, color = "black"),
        strip.text.y.right = element_text(angle = 0, size = 16)) +
  geom_hline(yintercept = 0.0, linewidth = 1) + 
  labs(fill = "")+ 
  guides(fill = guide_legend(nrow = 1))

#ggsave("figures/ms_figs/all_freq/all_freq_survey.png", width = 8, height = 6, units = ("in"))
#ggsave("figures/slides_svg/freq_survey.svg", width = 8, height = 6, units = ("in"))

#bridge freq
#reformat data
code_freq_s$code_cat <- factor(code_freq_s$code_cat, levels = c("Data bridge", "Data barrier", "Ecological bridge", "Ecological barrier", "Social bridge", "Social barrier", "Implementation bridge","Implementation barrier"))

code_sum_br_s <- code_freq_s %>%
  filter(bridge_barr == "Bridge") %>%
  #mutate(n_total = sum(total)) %>%
  #group_by(code_cat, code_name) %>%
  #summarise(value = total/n_total) %>% #percentages represent total number of bridge or barriers codes that code represents
  mutate(#percent = scales::percent(value, accuracy = 1), 
         code_name = factor(code_name)) %>%
  ungroup() %>%
  group_by(code_cat) %>%
  arrange(code_name, desc(total))

#plot
br_freq_s <- ggplot(code_sum_br_s, aes(x = total, y = reorder(code_name, total))) + 
  geom_point(size = 5, color = "lightskyblue3") + 
  # scale_x_continuous(limits = c(0, 30), 
  #                    breaks = seq(0, 30, 5), 
  #                    name = "Percent appearance (%)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  scale_x_continuous(limits = c(0, 20), 
                  #breaks = seq(0, 30, 5), 
                  name = "Sub-theme count (#)")+
  scale_y_discrete(position = "right") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 0), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

#ggsave("figures/ms_figs/all_freq/br_freq_survey.png", width = 6, height = 8, units = c("in"))

#barrier freq
#reformat data
code_sum_ba_s <- code_freq_s %>%
  filter(bridge_barr == "Barrier") %>%
  #mutate(n_total = sum(total)) %>%
  #group_by(code_cat, code_name) %>%
  #summarise(value = total/n_total) %>% #percentages represent total number of bridge or barriers codes that code represents
  mutate(#percent = scales::percent(value, accuracy = 1), 
         code_name = factor(code_name)) %>%
  ungroup() %>%
  group_by(code_cat) %>%
  arrange(code_name, desc(total))

#plot
ba_freq_s <- ggplot(code_sum_ba_s, aes(x = total, y = reorder(code_name, total))) + 
  geom_point(size = 5, color = "goldenrod") + 
  scale_x_reverse(limits = c(20, 0), 
                     breaks = seq(0, 20, 5), 
                     name = "Sub-theme count (#)") +
  theme_minimal() +
  facet_wrap("code_cat", ncol = 1, scales = "free_y") +
  theme(strip.placement = "outside",
        strip.text = element_text(face = "bold", size = 14, hjust = 1), 
        axis.text = element_text(size = 12, color = "black"), 
        axis.title = element_text(size = 14, color = "black")) +
  ylab("")

#ggsave("figures/ms_figs/all_freq/ba_freq_survey.png", width = 6, height = 8, units = c("in"))

```

![](images/br_ba_freq_survey.png){fig-align="center" width="10in", height = "12in"}
:::

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 5
#| fig-align: center
#| include: false

code_ts <- code_vars %>% 
  mutate(year = str_extract_all(Doc, "\\d+\\.?\\d*"), 
         year = as.Date(as.character(year), format = "%Y"), 
         year = year(year)) %>%
  select(c(Doc, year, Co.management.and.communication)) %>%
  group_by(year) %>%
  summarise(n_appear = sum(Co.management.and.communication), 
            n_year = n())

colnames(code_ts) <- c("Year", "com", "pubs")

ggplot(code_ts, aes(as.numeric(Year))) + 
  geom_path(aes(y = com), linewidth = 1.5, color = "dodgerblue4") + 
  geom_point(aes(y= com), size = 5, fill = "dodgerblue4", color = "black", shape = 21) + 
  geom_path(aes(y = pubs), linewidth = 1.5, color = "grey60") + 
  geom_point(aes(y = pubs), size = 5, fill = "grey60", color = "black", shape = 21) +
  xlab("") +
  ylab("Count (#)") +
  theme_minimal() + 
  theme(axis.text = element_text(size = 14, color = "black"), 
        axis.title = element_text(size = 16, color = "black")) 
```

### Result 2.2: Heat maps of theme co-occurence

This heatmap represents number of times two codes co-occurered together in the same article, or "theme co-occurence". Bridges are placed on rows and barriers are separated by columns. Literature codes are on the top and survey codes are on the bottom.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| include: FALSE

#literature
code_rel_melt <- reshape::melt(code_rel2)
code_rel_melt$variable <- gsub(".", " ", code_rel_melt$variable, fixed = TRUE)

code_rel_melt <- code_rel_melt %>%
  separate(Code.System, c("key", "code1"), "> ", convert = TRUE)

code_rel_melt$key <- str_trim(code_rel_melt$key)

colnames(code_rel_melt) <- c("brba_code1", "code1", "code2", "total")

#add brba column to second code col
code_rel_melt2 <- code_rel_melt
#code_rel_melt2$brba_code2 <- NA

code_rel_melt2$code2[code_rel_melt2$code2 == "Fine scale management appropriate for target species"] <- "Fine-scale management appropriate for target species"
code_rel_melt2$code2[code_rel_melt2$code2 == "Pre existing regulation and management institutions"] <- "Pre-existing regulation and management institutions"
code_rel_melt2$code2[code_rel_melt2$code2 == "Co management and communication"] <- "Co-management and communication"

# for(i in 1:nrow(code_rel_melt2)){
#   code_temp = code_rel_melt2$code2[i]
# 
#   brba_temp = code_rel_melt2 %>% filter(code1 == code_temp & code2 == code_temp)
#   brba_code_temp = brba_temp$brba_code1[1]
# 
#   code_rel_melt2$brba_code2[i] <- brba_code_temp
# }
# 
# code_rel_melt2$brba_code2[code_rel_melt2$code2 == "Target species is a prey specialist"] <- "Ecological bridge"

# group_colors = c("Data Bridge" = "dodgerblue4", "Data Barrier" = "dodgerblue4", "Ecological Bridge" = "honeydew4", "Ecological Barrier" = "honeydew4", "Social Bridge" = "darkorange4", "Social Barrier" = "darkorange4")

code_rel_melt_h <- code_rel_melt2 %>%
  #filter(total >= 11) %>%
  mutate(#bor_box = ifelse(total >= 11, "TRUE", "FALSE"), 
        #bor_box = as.factor(bor_box), 
        code1 = as.factor(code1), 
        code2 = as.factor(code2))

code_rel_melt_h$code1 <- factor(code_rel_melt_h$code1, levels = c(as.character(unique(code_rel_melt_h$code1))))

code_rel_melt_h$code2 <- factor(code_rel_melt_h$code2, levels =
c("Need for multispecies management",	"Implementation costs",	"Indirect socioeconomic consequences",	"Discontinuity",	"Poor communication and outreach",	"Poor program implementation", "Model abuse", "Gaps and inaccuracy", "Climate change adding uncertainty to target species habitat use",	"Management scales do not align with system needs"))

#code_rel_melt_h$bor_box <- factor(code_rel_melt_h$bor_box, levels = c("TRUE", "FALSE"))

# code_rel_melt_h <- code_rel_melt_h %>%
#   mutate(code1 = reorder(code1, -total, FUN = mean),
#          code2 = reorder(code2, total, FUN = mean))

# unique_x_groups <- code_rel_melt_h %>%
# distinct(code1, brba_code1) %>%
# arrange(code1)
# 
# unique_y_groups <- code_rel_melt_h %>%
# distinct(code2, brba_code2) %>%
# arrange(code2)

# x_colors <- setNames(group_colors[unique_x_groups$brba_code1], unique_x_groups$code1)
# y_colors <- setNames(group_colors[unique_y_groups$brba_code2], unique_y_groups$code2)

ggplot(code_rel_melt_h, aes(x = code2, y = code1)) +
geom_tile(aes(fill = total)) + 
scale_fill_gradient2(high = "red")+
ylab("")+
xlab("")+
scale_x_discrete(position = "top") +
labs(fill = "Number of co-occurences")+
theme_minimal() + 
theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.3, size = 13), 
      axis.text.y = element_text(size = 13))+
geom_text(aes(label = total))+
scale_color_manual(values = c("black", "#00000000"))+
guides(color = "none")

#ggsave("figures/ms_figs/heat_tile.png", width = 11, height = 12, units = c("in"))

#surveys
#reshape data
code_rel_melt_s <- reshape::melt(code_rel_s)
code_rel_melt_s$variable <- gsub(".", " ", code_rel_melt_s$variable, fixed = TRUE)

code_rel_melt_s <- code_rel_melt_s %>%
  separate(Code.System, c("key", "code1"), "> ", convert = TRUE)

code_rel_melt_s$key <- str_trim(code_rel_melt_s$key)

colnames(code_rel_melt_s) <- c("brba_code1", "code1", "code2", "total")

#add brba column to second code col
code_rel_melt2_s <- code_rel_melt_s
#code_rel_melt2$brba_code2 <- NA

code_rel_melt2_s$code2[code_rel_melt2_s$code2 == "Interest holder conflicts"] <- "Interest-holder conflicts"
code_rel_melt2_s$code2[code_rel_melt2_s$code2 == "Missing cross program comparisons"] <- "Missing cross-program comparisons"
code_rel_melt2_s$code2[code_rel_melt2_s$code2 == "Lack of near real time data"] <- "Lack of near real-time data"

code_rel_melt_h_s <- code_rel_melt2_s %>%
  #filter(total >= 11) %>%
  mutate(#bor_box = ifelse(total >= 11, "TRUE", "FALSE"), 
        #bor_box = as.factor(bor_box), 
        code1 = as.factor(code1), 
        code2 = as.factor(code2))

#relevel
code_rel_melt_h_s$code1 <- factor(code_rel_melt_h_s$code1, levels = c(as.character(unique(code_rel_melt_h_s$code1))))

code_rel_melt_h_s$code2 <- factor(code_rel_melt_h_s$code2, levels =
c("Unpredictable bycatch patterns", "Missing cross-program comparisons",	"Gaps and inaccuracy",	"Lack of near real-time data", "Cost",	"Discontinuity",	"Confidentiality",	"Lack of incentives",	"Volunteer programs reduce participation",	"Regulated programs reduce participation",	"External factors reducing participation",	"Poor communication and outreach",	"Interest-holder conflicts",	"Stigmatization", "Difficulty assessing impact",	"Limited relevance"))

#plot
ggplot(code_rel_melt_h_s, aes(x = code2, y = code1)) +
geom_tile(aes(fill = total)) + 
scale_fill_gradient2(high = "red")+
ylab("")+
xlab("")+
scale_x_discrete(position = "top") +
labs(fill = "Number of co-occurences")+
theme_minimal() + 
theme(axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.3, size = 13), 
      axis.text.y = element_text(size = 13))+
geom_text(aes(label = total))+
scale_color_manual(values = c("black", "#00000000"))+
guides(color = "none")

#ggsave("figures/ms_figs/heat_tile_survey.png", width = 8.5, height = 8, units = c("in"))

```

![](images/heat_tile_ai.png){fig-align="center" width="10in", height = "12in"}

![](images/heat_tile_survey_ai.png){fig-align="center" width="10in", height = "12in"}

## Result 3: Variations by industry

### 3.1: Management and conservation problems DOM has been used to address

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 4
#| fig-align: center

code_vars$broad_location <- as.factor(code_vars$broad_location)
codes_LU <- code_vars[!duplicated(code_vars$Lit_unit), ]

consv <- codes_LU %>%
  gather(col, consv_topic, reason1:reason3) %>%
  select(consv_topic)

consv_s <- consv %>%
  group_by(consv_topic) %>%
  filter(consv_topic != "NA") %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(perc = count/sum(count)*100)

consv_bar <- ggplot(consv_s) +
  geom_bar(aes(x = perc, y = reorder(consv_topic, perc)), fill = "dodgerblue4", width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     expand = c(0,0), 
                     limits = c(0, 45))+
  scale_y_discrete(expand = expansion(add = c(0,0))) + 
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Arial", size = 12)) + 
  geom_text(aes(x = perc, y = consv_topic, label = consv_topic),
    hjust = 0,
    nudge_x = 0.3,
    colour = "dodgerblue4",
    family = "Arial",
    size = 3.5)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Arial",
      size = 13)
  )

consv_bar
```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 9
#| fig-height: 12
#| fig-align: center
consv <- codes_LU %>%
  gather(col, consv_topic, reason1:reason3) %>%
  gather(test, fishery_channel, fishery_or_channel1:fishery_or_channel4) %>%
  select(c(consv_topic, fishery_channel)) %>%
  mutate(consv_bin = paste0(consv_topic, " - ", fishery_channel))

consv_s <- consv %>%
  group_by(consv_topic, consv_bin) %>%
  filter(consv_topic != "NA") %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(perc = count/sum(count)*100)

consv_bar2 <- ggplot(consv_s) +
  geom_bar(aes(x = perc, y = reorder(consv_bin, perc), fill = consv_topic), width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     expand = c(0,0), 
                     limits = c(0, 45))+
  scale_y_discrete(expand = expansion(add = c(0,0))) + 
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Arial", size = 12)) + 
  geom_text(aes(x = perc, y = consv_bin, label = consv_bin),
    hjust = 0,
    nudge_x = 0.3,
    colour = "dodgerblue4",
    family = "Arial",
    size = 3.5)+
  labs(subtitle = "Percent of sum from all articles (%)") + 
  theme(plot.subtitle = element_text(
      family = "Arial",
      size = 13)
  )

consv_bar2
```

### 3.2: Fishing vs. shipping

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center
#| include: false

#survey figs

#format data
freq_job_survey <- code_vars_s %>% 
  select(c(1, 21:54)) %>% 
  janitor::clean_names() %>%
  pivot_longer(2:35, names_to = "sub_theme", values_to = "total") %>%
  group_by(survey, sub_theme) %>%
  summarise(total = sum(total), 
            .groups = "drop") %>%
  mutate(fish_ship = ifelse(str_detect(survey, "WhaleWatch"), "Shipping", "Fishing"))

lib_job_survey <- code_freq_s %>%
  mutate(id = paste(bridge_barr, code_cat, code_name, sep = "_")) %>% 
  select(id)

freq_job_survey$id <- NA
for(i in 1:nrow(freq_job_survey)){
  curr_code = freq_job_survey$sub_theme[i]
  curr_id = lib_job_survey$id[agrepl(curr_code, lib_job_survey$id, 6)]

  freq_job_survey$id[i] <- curr_id

}

freq_job_survey$id[freq_job_survey$sub_theme == "stigmatization"] <- "Barrier_Social barrier_Stigmatization"
freq_job_survey$id[freq_job_survey$sub_theme == "cost"] <- "Barrier_Social barrier_Cost"

freq_job_survey <- freq_job_survey %>%
  separate_wider_delim(id, delim = "_", names = c("bridge_barr", "code_cat", "code_name"))

freq_job_survey$bridge_barr <- factor(freq_job_survey$bridge_barr, levels = c("Bridge", "Barrier"))

#calculate proportions
freq_job_survey2 <- freq_job_survey %>%
  group_by(fish_ship, bridge_barr) %>%
  mutate(n_total = sum(total)) %>%
  ungroup() %>%
  group_by(fish_ship, bridge_barr, code_cat) %>% 
  mutate(n_subtotal = sum(total)) %>%
  reframe(value = (n_subtotal/n_total)*100) %>%
  ungroup() %>%
  mutate(code_level = if_else(code_cat %in% c("Ecological barrier", "Ecological bridge"), "Ecological bridge or barrier", 
                              if_else(code_cat %in% c("Data barrier", "Data bridge"), "Data bridge or barrier",
                              if_else(code_cat %in% c("Social barrier", "Social bridge"), "Social bridge or barrier", 
                              if_else(code_cat %in% c("Implementation bridge", "Implementation barrier"), "Implementation bridge or barrier", NA)))), 
          id = paste(fish_ship, bridge_barr, code_cat)) %>%
  distinct(id, .keep_all = TRUE)

#stacked bar plots
ggplot(freq_job_survey2, aes(fill=code_level, y=value, x=fish_ship)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual(values = c("#CFE2F3", "#C5E0B4","royalblue4","#45818E")) +
    facet_wrap(~bridge_barr) +
    labs(fill = "") +
    guides(fill=guide_legend(nrow=2,byrow=TRUE))+
    theme_ms() +
    theme(legend.position = "top", 
          legend.justification = "left", 
          legend.text = element_text(size = 14, color = "black"))+
    xlab("") + 
    ylab("Proportion of theme (%)")

#ggsave("figures/ms_figs/industry/stack_survey.png", width = 7, height = 6, units = c("in"))
#ggsave("figures/slides_svg/stack_survey.svg", width = 7, height = 6, units = c("in"))

#radar plots
freq_job_survey3 <- freq_job_survey %>%
  # group_by(fish_ship, bridge_barr) %>%
  # mutate(n_total = sum(total)) %>%
  # ungroup() %>%
  # group_by(fish_ship, bridge_barr, code_name) %>%
  # mutate(n_subtotal = sum(total)) %>%
  # reframe(value = (n_subtotal/n_total)*100) %>%
  # ungroup() %>%
  #mutate(id = paste(fish_ship, bridge_barr, code_name)) %>%
  #distinct(id, .keep_all = TRUE) %>%
  #select(-c(id))
  group_by(fish_ship, bridge_barr, code_name) %>%
  summarise(value = sum(total))

top_survey <- freq_job_survey3 %>% 
                group_by(fish_ship, bridge_barr) %>% 
                arrange(desc(value)) %>% 
                top_n(3) %>% 
                ungroup() %>%
                select(code_name)

top_bridge <- freq_job_survey3 %>%
  ungroup() %>%
  subset(code_name %in% top_survey$code_name) %>% 
  filter(bridge_barr == "Bridge") %>%
  pivot_wider(names_from = code_name, values_from = value) %>% 
  select(-bridge_barr) %>%
  mutate_all(~replace_na(., 0))

top_barrier <- freq_job_survey3 %>%
  ungroup() %>%
  subset(code_name %in% top_survey$code_name) %>%
  filter(bridge_barr == "Barrier") %>% 
  pivot_wider(names_from = code_name, values_from = value) %>% 
  select(-bridge_barr) %>%
  mutate_all(~replace_na(., 0))

ggradar::ggradar(top_bridge,
        values.radar = c(0, 5, 10), 
        background.circle.colour = "white",
        axis.line.colour = "gray60",
        gridline.min.colour = "gray60",
        gridline.mid.colour = "gray60",
        gridline.max.colour = "gray60",
        group.colours = c("#F29325", "#007172"),#007172
        legend.position = "bottom", 
        fill = TRUE, 
        axis.label.size = 3)

#ggsave("figures/ms_figs/industry/radar_br_survey.png", height = 8, width = 8, units = c("in"))
#ggsave("figures/slides_svg/radar_br.svg", height = 8, width = 8, units = c("in"))

ggradar::ggradar(top_barrier,
        values.radar = c(0, 3, 6), 
        background.circle.colour = "white",
        axis.line.colour = "gray60",
        gridline.min.colour = "gray60",
        gridline.mid.colour = "gray60",
        gridline.max.colour = "gray60",
        group.colours = c("#F29325", "#007172"),#007172
        legend.position = "none", 
        fill = TRUE, 
        axis.label.size = 3)

#ggsave("figures/ms_figs/industry/radar_ba_survey.png", height = 8, width = 8, units = c("in"))
#ggsave("figures/slides_svg/radar_ba.svg", height = 8, width = 8, units = c("in"))

#get sample sizes of survey respondants by position
n_job_survey <- code_vars_s %>% 
  select(c(1, 21:54)) %>% 
  janitor::clean_names()

table(n_job_survey$survey)

```

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center
#| include: false

#literature figs

#format data
freq_job_lit <- code_vars %>% 
  select(c(1, 27:29, 33:60)) %>%
  janitor::clean_names() %>%
  pivot_longer(2:4, names_to = "test", values_to = "fish_ship") %>%
  select(-c(test)) %>%
  pivot_longer(2:29, names_to = "sub_theme", values_to = "total") %>%
  group_by(fish_ship, sub_theme) %>%
  summarise(total = sum(total), 
            .groups = "drop") %>%
  filter(fish_ship == "Bycatch" | fish_ship == "Ship strikes" | fish_ship == "Overfishing" | fish_ship == "Entanglement" | fish_ship == "Avoid fisheries discharge")

lib_job_lit <- code_freq %>%
  mutate(id = paste(bridge_barr, code_cat, code_name, sep = "_")) %>% 
  select(id)

freq_job_lit$id <- NA
for(i in 1:nrow(freq_job_lit)){
  curr_code = freq_job_lit$sub_theme[i]
  curr_id = lib_job_lit$id[agrepl(curr_code, lib_job_lit$id, 10)] #6 refers to the number of characters that can be diff

  freq_job_lit$id[i] <- curr_id

}

freq_job_lit$id[freq_job_lit$sub_theme == "abiotic_data_availability"] <- "Bridge_Data bridge_Abiotic data availability"
freq_job_lit$id[freq_job_lit$sub_theme == "abiotic_indicators_of_target_species_presence"] <- 
"Bridge_Ecological bridge_Abiotic indicators of target species presence"
freq_job_lit$id[freq_job_lit$sub_theme == "cost"] <- 
"Barrier_Social barrier_Cost"
freq_job_lit$id[freq_job_lit$sub_theme == "discontinuity"] <- 
"Barrier_Social barrier_Discontinuity"
freq_job_lit$id[freq_job_lit$sub_theme == "feasibility"] <- 
"Bridge_Social bridge_Feasibility"
freq_job_lit$id[freq_job_lit$sub_theme == "model_abuse"] <- 
"Barrier_Data barrier_Model abuse"
freq_job_lit$id[freq_job_lit$sub_theme == "resource_availability"] <- 
"Bridge_Social bridge_Resource availability"

freq_job_lit <- freq_job_lit %>%
  separate_wider_delim(id, delim = "_", names = c("bridge_barr", "code_cat", "code_name"))

freq_job_lit$bridge_barr <- factor(freq_job_lit$bridge_barr, levels = c("Bridge", "Barrier"))

freq_job_lit <- freq_job_lit %>%
  mutate(fish_ship = ifelse(fish_ship == "Ship strikes", "Shipping", "Fishing"))

#calculate proportions
freq_job_lit2 <- freq_job_lit %>%
  group_by(fish_ship, bridge_barr) %>%
  mutate(n_total = sum(total)) %>%
  ungroup() %>%
  group_by(fish_ship, bridge_barr, code_cat) %>% 
  mutate(n_subtotal = sum(total)) %>%
  reframe(value = (n_subtotal/n_total)*100) %>%
  ungroup() %>%
  mutate(code_level = if_else(code_cat %in% c("Ecological barrier", "Ecological bridge"), "Ecological bridge or barrier", 
                              if_else(code_cat %in% c("Data barrier", "Data bridge"), "Data bridge or barrier",
                              if_else(code_cat %in% c("Social barrier", "Social bridge"), "Social bridge or barrier", NA))), 
          id = paste(fish_ship, bridge_barr, code_cat)) %>%
  distinct(id, .keep_all = TRUE)

#stacked bar plots
ggplot(freq_job_lit2, aes(fill=code_level, y=value, x=fish_ship)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual(values = c("#CFE2F3", "#C5E0B4","#45818E")) +
    facet_wrap(~bridge_barr) +
    labs(fill = "") +
    guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
    theme_ms() +
    theme(legend.position = "top", 
          legend.justification = "left", 
          legend.text = element_text(size = 14, color = "black")) +
    xlab("") + 
    ylab("Proportion of theme (%)")

#ggsave("figures/ms_figs/industry/stack_lit.png", width = 7, height = 6, units = c("in"))

#radar plots
freq_job_lit3 <- freq_job_lit %>%
  # group_by(fish_ship, bridge_barr) %>%
  # mutate(n_total = sum(total)) %>%
  # ungroup() %>%
  # group_by(fish_ship, bridge_barr, code_name) %>% 
  # mutate(n_subtotal = sum(total)) %>%
  # reframe(value = (n_subtotal/n_total)*100) %>%
  # ungroup() %>%
  # mutate(id = paste(fish_ship, bridge_barr, code_name)) %>%
  # distinct(id, .keep_all = TRUE) %>%
  # select(-c(id)) %>%
  group_by(fish_ship, bridge_barr, code_name) %>%
  summarise(value = sum(total))

top_lit <- freq_job_lit3 %>% 
                group_by(fish_ship, bridge_barr) %>% 
                arrange(desc(value)) %>% 
                top_n(3) %>% 
                ungroup() %>%
                select(code_name)

top_bridge <- freq_job_lit3 %>%
  subset(code_name %in% top_lit$code_name) %>% 
  filter(bridge_barr == "Bridge") %>% 
  pivot_wider(names_from = code_name, values_from = value) %>% 
  ungroup() %>%
  select(-bridge_barr) %>%
  mutate_all(~replace_na(., 0))

top_barrier <- freq_job_lit3 %>%
  subset(code_name %in% top_lit$code_name) %>%
  filter(bridge_barr == "Barrier") %>% 
  pivot_wider(names_from = code_name, values_from = value) %>% 
  ungroup() %>%
  select(-bridge_barr) %>%
  mutate_all(~replace_na(., 0))

ggradar::ggradar(top_bridge,
        values.radar = c(0, 20, 40), 
        background.circle.colour = "white",
        axis.line.colour = "gray60",
        gridline.min.colour = "gray60",
        gridline.mid.colour = "gray60",
        gridline.max.colour = "gray60",
        group.colours = c("#F29325", "#007172"),
        legend.position = "bottom", 
        fill = TRUE, 
        axis.label.size = 3)

#ggsave("figures/ms_figs/industry/radar_br_lit.png", height = 8, width = 8, units = c("in"))

ggradar::ggradar(top_barrier,
        values.radar = c(0, 15, 35), 
        background.circle.colour = "white",
        axis.line.colour = "gray60",
        gridline.min.colour = "gray60",
        gridline.mid.colour = "gray60",
        gridline.max.colour = "gray60",
        group.colours = c("#F29325", "#007172"),
        legend.position = "none", 
        fill = TRUE, 
        axis.label.size = 3)

#("figures/ms_figs/industry/radar_ba_lit.png", height = 8, width = 8, units = c("in"))

#get sample sizes of survey respondants by position
n_job_lit <- code_vars %>% 
  select(c(1, 27:29, 33:60)) %>%
  janitor::clean_names()  %>%
  pivot_longer(2:4, names_to = "test", values_to = "fish_ship") %>%
  select(-c(test)) %>%
  filter(fish_ship == "Bycatch" | fish_ship == "Ship strikes" | fish_ship == "Overfishing" | fish_ship == "Entanglement" | fish_ship == "Avoid fisheries discharge") %>%
  mutate(fish_ship = ifelse(fish_ship == "Ship strikes", "Shipping", "Fishing"))

table(n_job_lit$fish_ship)

```

::: panel-tabset

#### Literature

![](images/lit_combo.png){fig-align="center" width="5in", height = "8in"}

#### Surveys

![](images/survey_combo.png){fig-align="center" width="5in", height = "8in"}

:::

## Result 4: Variations by position (survey only)

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center
#| include: false

#survey figs

#format data
freq_pos_survey <- code_vars_s %>% 
  select(c(6:10, 21:54)) %>% 
  janitor::clean_names() %>%
  pivot_longer(1:5, names_to = "test", values_to = "position") %>%
  select(-c(test)) %>%
  pivot_longer(1:34, names_to = "sub_theme", values_to = "total") %>%
  group_by(position, sub_theme) %>%
  summarise(total = sum(total), 
            .groups = "drop") %>%
  drop_na(position)

lib_job_survey <- code_freq_s %>%
  mutate(id = paste(bridge_barr, code_cat, code_name, sep = "_")) %>% 
  select(id)

freq_pos_survey$id <- NA
for(i in 1:nrow(freq_pos_survey)){
  curr_code = freq_pos_survey$sub_theme[i]
  curr_id = lib_job_survey$id[agrepl(curr_code, lib_job_survey$id, 6)]

  freq_pos_survey$id[i] <- curr_id

}

freq_pos_survey$id[freq_pos_survey$sub_theme == "stigmatization"] <- "Barrier_Social barrier_Stigmatization"
freq_pos_survey$id[freq_pos_survey$sub_theme == "cost"] <- "Barrier_Social barrier_Cost"

freq_pos_survey <- freq_pos_survey %>%
  separate_wider_delim(id, delim = "_", names = c("bridge_barr", "code_cat", "code_name"))

freq_pos_survey$bridge_barr <- factor(freq_pos_survey$bridge_barr, levels = c("Bridge", "Barrier"))

#calculate proportions
freq_pos_survey2 <- freq_pos_survey %>%
  group_by(position, bridge_barr) %>%
  mutate(n_total = sum(total)) %>%
  ungroup() %>%
  group_by(position, bridge_barr, code_cat) %>% 
  mutate(n_subtotal = sum(total)) %>%
  reframe(value = (n_subtotal/n_total)*100) %>%
  ungroup() %>%
  mutate(code_level = if_else(code_cat %in% c("Ecological barrier", "Ecological bridge"), "Ecological bridge or barrier", 
                              if_else(code_cat %in% c("Data barrier", "Data bridge"), "Data bridge or barrier",
                              if_else(code_cat %in% c("Social barrier", "Social bridge"), "Social bridge or barrier", 
                              if_else(code_cat %in% c("Implementation bridge", "Implementation barrier"), "Implementation bridge or barrier", NA)))), 
          id = paste(position, bridge_barr, code_cat)) %>%
  distinct(id, .keep_all = TRUE)

#stacked bar plots
ggplot(freq_pos_survey2, aes(fill=code_level, y=value, x=position)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual(values = c("#CFE2F3", "#C5E0B4","royalblue4","#45818E")) +
    facet_wrap(~bridge_barr) +
    labs(fill = "") +
    theme_ms() +
    theme(legend.position = "top", 
          legend.justification.top = "left", 
          legend.title = element_text(size = 16, color = "black"), 
          legend.text = element_text(size = 14, color = "black"))+
    guides(fill=guide_legend(nrow=2,byrow=TRUE))+
    xlab("") + 
    ylab("Proportion of theme (%)")+
    theme(axis.text.x = element_text(size = 14, color = "black", angle = 60, hjust = 1))

#ggsave("figures/ms_figs/position/stack_survey.png", width = 10, height = 8, units = c("in"))

#get sample sizes of survey respondants by position
n_pos_survey <- code_vars_s %>% 
  select(c(6:10, 21:54)) %>% 
  janitor::clean_names() %>%
  pivot_longer(1:5, names_to = "test", values_to = "position") %>%
  select(-c(test)) %>%
  drop_na(position)

table(n_pos_survey$position)

```

![](images/position.png){fig-align="center" width="8in", height = "6in"}